Algebraic filtering of surfaces from medical images with Julia

### 2019-06-30 src/background.tex
my last update

A \emph{representation scheme} for solid modeling is a mapping between a space of mathematical models and a space of symbolic representations, often generated by a formal grammar.
Solid pointsets (i.e., `$r$-sets') are defined~\cite{requicha:80} as compact (bounded and closed) regular and semianalytic subsets of the $d$-space. A large number of representation schemes were defined in the past forty years, including the two main classes of (a) \emph{boundary representations} (`$B$-reps'), where the solid model is represented through a representation of its boundary elements, i.e.~faces, edges and vertices, and (b) \emph{decompositive/enumerative representations}, that are a decomposition of either the object or the embedding space, respectively, into a well-defined cellular complex. In particular, a boundary representation provides a cellular decomposition of the object's boundary into cells of dimension zero (vertices), one (edges), and two (faces). Medical imaging can be classified as \emph{enumerative representation} of cellular decompositions of organs and tissues of interest, in particular, as subsets \emph{of 3D volume elements} (voxels) from the 3D image.

@@

The \emph{Linear Algebraic Representation} (\textsc{lar}) aims to represent the \emph{chain complex} generated by a piecewise-linear geometric complex embedded either in 2D or in 3D. In few words, it gets a minimal characterization of geometry and topology of a cellular complex, i.e.~the embedding mapping $\mu : C_0 \to \E^d$ of 0-cells (vertices), as well a description of $(d-1)$-cells as subsets of vertices, and is able to return the whole chain complex 
\[ 
C_\bullet = (C_p, \partial_p) := 
C_3 \ 
\substack{
\delta_2 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_3 
}
\ C_2 \ 
\substack{
\delta_1 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_2 
}
\ C_1 \ 
\substack{
\delta_0 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_1 
}
\ C_0 .
\] 
and, in particular, any basis for linear chain spaces $C_p$, and any linear
boundary/coboundary map \(\partial_p\) and
\(\delta_p=\partial_{p+1}^\top\) between them. The domain of \textsc{lar} is the set of chain complexes generated by cell $d$-complexes ($2\leq d\leq 3$). The computer representations of \textsc{lar} are sparse binary matrices to represent both the operators and the chain bases. Note that in algebraic topology a $p$-chain is defined as a linear combination of $p$-cells with scalars from a field. When the scalar coefficients are from $\{-1, 0, +1\}$, a chain may represent \emph{any (oriented) subset of cells} from the cellular complex.

>>>>

A \emph{representation scheme} for solid modeling is a mapping between a space of mathematical models and a space of symbolic representations, often generated by a formal grammar.
Solid pointsets (i.e., `$r$-sets') are defined~\cite{Requicha:1980:RRS:356827.356833} as compact (bounded and closed) regular and semianalytic subsets of the $d$-space. A large number of representation schemes were defined in the past forty years, including the two main classes of (a) \emph{boundary representations} (`$B$-reps'), where the solid model is represented through a representation of its boundary elements, i.e.~faces, edges and vertices, and (b) \emph{decompositive/enumerative representations}, that are a decomposition of either the object or the embedding space, respectively, into a well-defined \emph{cellular complex}. In particular, a boundary representation provides a cellular decomposition of the object's boundary into \emph{cells} of dimension zero (vertices), one (edges), and two (faces). Medical imaging can be classified as \emph{enumerative representation} of cellular decompositions of organs and tissues of interest, in particular, as subsets \emph{of 3D volume elements} (voxels) from the 3D image.

@@

The \emph{Linear Algebraic Representation} (\textsc{lar})~\cite{Dicarlo:2014:TNL:2543138.2543294} aims to represent the \emph{chain complex}~\cite{TSAS} generated by a piecewise-linear geometric complex embedded either in 2D or in 3D. In few words, it gets a minimal characterization of geometry and topology of a cellular complex, i.e.~the embedding mapping $\mu : C_0 \to \E^d$ of 0-cells (vertices), as well a description of $(d-1)$-cells as subsets of vertices, and is able to return the whole chain complex 
\[ 
C_\bullet = (C_p, \partial_p) := 
C_3 \ 
\substack{
\delta_2 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_3 
}
\ C_2 \ 
\substack{
\delta_1 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_2 
}
\ C_1 \ 
\substack{
\delta_0 \\
\longleftarrow \\[-1mm]
\longrightarrow \\
\partial_1 
}
\ C_0 .
\] 
and, in particular, any basis for linear chain spaces $C_p$, and any linear
boundary/coboundary map \(\partial_p\) and
\(\delta_p=\partial_{p+1}^\top\) between them. The domain of \textsc{lar} is the set of chain complexes generated by cell $d$-complexes ($2\leq d\leq 3$). The computer representations of \textsc{lar} are sparse binary matrices to represent both the operators and the chain bases. Note that in algebraic topology a $p$-chain is defined as a linear combination of $p$-cells with scalars from a field. When the scalar coefficients are from $\{-1, 0, +1\}$, a chain may represent \emph{any (oriented) subset of cells} from the cellular complex.

### 2019-06-30 src/design.tex
posted contents for implementation section
my last update

For the sake of simplicity, we assume a common size on the three image axes, and the corresponding image portion $B$, called \emph{block}, as a function of its element of the  lowest  block  coordinates $i,j,k\in\N$ and of block dimension $n$:
\[
B(i,j,k,n) := Image([in\mbox{:}in+n, jn\mbox{:}jn+n, kn\mbox{:}kn+n]) 
\]

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.5\linewidth]{figs/blocks} 
   \caption{A possible block partitioning of a radiologic image. The evidenced 2D block, of size $n^d=64^2$, is sliced as $B([2,1,64]) = Image([128:172],[64:128)]$}
   \label{fig:blocks}
\end{figure}

@@

\paragraph{Chain coordinates }\label{sec:bbbb} 
We are going to treat each image block independently from each other. Hence we map each image subspace $B(i,j,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n$, using coordinate vectors $c_{h,k}\in \B^{n\times n} := \{0,1\}^{n\times n}$, where the basis element $c = c_{h,k} \in C_2$ is mapped via Cartesian-to-linear map to the Boolean vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

Therefore, each pixel (or voxel) in a block image will be seen as a basis Boolean vector in $C_2$, and each subset of image elements, as the corresponding Boolean vector in $C_2$, with many ones as the cardinality of the subset.

@@

@@

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. We aim to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the block, in order to decompose the input $Image(u,v,w)$ into a fair number 
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n} = \frac{uvw}{n^3}
\] of blocks. 
Then we consider each image portion $c_{i,j,k} = S\cap B(i,j,k,n)$ and compute its (binary) coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$. Finally compute a matrix $\T{B}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{B} = [\partial_3(n)]\, \T{S}.
\]
where the boundary matrix has dimension $d\,n\,(1+n)^{d-1} \times n^d$.
Of course, the $\T{B}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S\cap Box_{i,j,k}{i,j,k}$, and the number of rows of the operator, equal to the dimension of the linear space $C_2$.

\paragraph{Embedding}
A final computational step is needed, in order to embed the 2-chains in Euclidean space $\E^3$ and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one corresponding to the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of \T{B}.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u$, $0\leq y\leq v$, and $0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, that will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

@@

All boundary surface subsets $B_{i,j,k}(S) = \partial_3 S \cap \mbox{Box}_{i,j,k}$, provided by  columns of $\T{S}$, are embedded in the same coordinate space. In formal terms: 
\[
\texttt{Lar}(S) := (\texttt{Geom}(S), \texttt{Top}(S)) = (\texttt{V}, \texttt{CV}),
\]
where, with respect to the \emph{chain complex} $C_3\to C_2\to C_1\to C_0$ induced by the input image $Im$ and segment $S_{i,j,k}$, we get
\begin{align}
\texttt{Geom} &:= \mu(C_0(_{i,j,k})) = \texttt{V},
\\
\texttt{Top} &:= C_3(S) = \T{S} \mapsto \texttt{CV}.
\end{align}
and
\begin{align}
\texttt{Lar}(B_{i,j,k}) &:= (\texttt{Geom}, \texttt{Top}) = (\texttt{W}, \texttt{FW}),
\\
\texttt{Geom} &:= \mu(C_0(B_{i,j,k})) = \texttt{W} \subset \texttt{V},
\\
\texttt{Top} &:= C_2(B_{i,j,k}) = \T{B}_{i,j,k} \mapsto \texttt{FW} \subset \texttt{FV}.
\end{align}

A translation transformation applied to each vertex subset $\texttt{W}_{i,j,k}$ with translation  vector $\v{t} = [i,j,k]$ will therefore move it in the final space position, so finally giving
\[
\texttt{Lar}(B) = \oplus_{i,j,k}\texttt{Lar}(\partial_3 S_{i,j,k})) = \oplus_{i,j,k}(\texttt{W}, \texttt{FW}) .
\]

>>>>

For the sake of simplicity, we assume a common size on the three image axes, and the corresponding image portion $\mathcal{B}$, called \emph{block}, as a function of its element of the  lowest  block  coordinates $i,j,k\in\N$ and of block dimension $n$:
\[
\mathcal{B}(i,j,k,n) := Image([in\mbox{:}in+n, jn\mbox{:}jn+n, kn\mbox{:}kn+n]) 
\]

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.5\linewidth]{figs/blocks} 
   \caption{A possible block partitioning of a radiologic image. The evidenced 2D block, of size $n^d=64^2$, is sliced as $\mathcal{B}([2,1,64]) = Image([128:172],[64:128)]$}
   \label{fig:blocks}
\end{figure}

@@

\paragraph{Chain coordinates }\label{sec:bbbb} 
We are going to treat each image block independently from each other. Hence we map each image subspace $\mathcal{B}(i,j,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n$, using coordinate vectors $c_{h,k}\in \\mathcal{B}^{n\times n} := \{0,1\}^{n\times n}$, where the basis element $c = c_{h,k} \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \\mathcal{B}^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

Therefore, each pixel (or voxel) in a block image will be seen as a basis binary vector in $C_2$, and each subset of image elements, as the corresponding binary vector in $C_2$, with many ones as the cardinality of the subset.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.4\linewidth]{figs/grid.pdf} 
   \caption{example caption}
   \label{fig:example}
\end{figure}

@@

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. We aim to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the block, in order to decompose the input $Image(u,v,w)$ into a fair number 
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n} \simeq \frac{uvw}{n^3}
\] of blocks. 
Then we consider each image portion $c_{i,j,k} = S\cap \mathcal{B}(i,j,k,n)$ and compute its (binary) coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$. Finally compute a matrix $\T{\mathcal{B}}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{\mathcal{B}} = [\partial_3(n)]\, \T{S}.
\]
where the boundary matrix has dimension $d\,n\,(1+n)^{d-1} \times n^d$.
Of course, the $\T{\mathcal{B}}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S\cap Box_{i,j,k}{i,j,k}$, and the number of rows of the operator, equal to the dimension of the linear space $C_2$.

\paragraph{Embedding}
A final computational step is needed, in order to embed the 2-chains in Euclidean space $\E^3$ and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one corresponding to the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of $\T{\mathcal{B}}$.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u$, $0\leq y\leq v$, and $0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, that will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

@@

All boundary surface subsets $\mathcal{B}_{i,j,k}(S) = \partial_3 S \cap \mbox{Box}_{i,j,k}$, provided by  columns of $\T{S}$, are embedded in the same coordinate space. In formal terms: 
\[
\texttt{Lar}(S) := (\texttt{Geom}(S), \texttt{Top}(S)) = (\texttt{V}, \texttt{CV}),
\]
where, with respect to the \emph{chain complex} $C_3\to C_2\to C_1\to C_0$ induced by the input image $Im$ and segment $S_{i,j,k}$, we get
\begin{align}
\texttt{Geom} &:= \mu(C_0(_{i,j,k})) = \texttt{V},
\\
\texttt{Top} &:= C_3(S) = \T{S} \mapsto \texttt{CV}.
\end{align}
and
\begin{align}
\texttt{Lar}(B_{i,j,k}) &:= (\texttt{Geom}, \texttt{Top}) = (\texttt{W}, \texttt{FW}),
\\
\texttt{Geom} &:= \mu(C_0(B_{i,j,k})) = \texttt{W} \subset \texttt{V},
\\
\texttt{Top} &:= C_2(B_{i,j,k}) = \T{\mathcal{B}}_{i,j,k} \mapsto \texttt{FW} \subset \texttt{FV}.
\end{align}

A translation transformation applied to each vertex subset $\texttt{W}_{i,j,k}$ with translation  vector $\v{t} = [i,j,k]$ will therefore move it in the final space position, so finally giving
\[
\texttt{Lar}(\mathcal{B}) = \oplus_{i,j,k}\texttt{Lar}(\partial_3 S_{i,j,k})) = \oplus_{i,j,k}(\texttt{W}, \texttt{FW}) .
\]

### 2019-06-30 src/implementation.tex
ircad experiment added
my last update

\subsection{Julia implementation}\label{sec:implementation}

@@

@@

\item depending on the above, best decision for $\emph{size}$ of 3D image blocks $B$ (bricks); defaults to: $\emph{size}=64$; hence the number of bricks will be $n=\ceil{\ell_1/size}\times\ceil{\ell_2/size}\times\ceil{\ell_3/size}$. 
Hence default value: $n=256$, for images $\mathcal{I} = 512\times 512\times 256$;

\item computation of a Julia's sparse boundary matrix $[\partial_B]$, returning a value of type \texttt{SparseMatrixCSC\{Int8\}\{Int64\}}, where \texttt{Int8} and \texttt{Int64} are the types for values and indices, respectively, stored by Compressed Sparse Column (CSC) format; the average $[\partial_B]$ (for $\emph{size}$ = 64) is about

\item creation of a either a local or distributed channel to implement a producer/consumer model of parallel/distributed computation, depending on available resources;

\item distribution of matrix $[\partial_B]$, of size $45$ MB, to all available nodes/cores (workers) using the macro \texttt{@eveywhere}.

@@

With $size=64$, the number of non-zeros within the sparse matrix $[\partial_{64^3}]$ is $\mathtt{nnz} = 4 792 266$, for a memory size of $9\times \mathtt{nnz}+8\times 262144 \simeq 45$ MB. The memory size of the sparse matrix is computed by considering $8+1$ bytes for non-zero element (which are exactly 6 per row), plus 8 bytes per each index of column start.

@@

\paragraph{3-Chain encoding} 
The interesting part of the \emph{Image} $\mathcal{I}$ is called \emph{Segment} $\mathcal{S}$. The goal of the whole \emph{workflow} is to extract a \emph{boundary model} of $\mathcal{S}$ from $\mathcal{I}$. The portion of $\mathcal{S}$ inside $B$, will be denoted as $\mathcal{S}(B)$.
Each block $B$ of the 3D image must by converted into the \emph{coordinate representation} of a vector $\nu\in C_3$ in the linear space  of 3-chains.

In coordinates local to $B$, once fixed an ordering from Cartesian to linear coords, this vector is represented by a \emph{binary array} of length $size^3$. With $size=64$, we have $64^3=262144$,  with a non-zero value (i.e.~$1$) for each foreground voxel in $\mathcal{S}(B)$. Therefore, the coded segment portion $\mathcal{S}(B)$, results with a space occupancy of about $262$ KB if encoded as a full array (i.e.~including the zero values). Whether encoded as a sparse vector, its space occupancy will correspondingly decrease.

@@

\item conversion from linear indices to Cartesian indices in coordinates local to the $B$ block, using the appropriate library functions;

@@

\paragraph{Assembling and artifact filtering} 
The results of the previous stages can be described as a \emph{collection of sets} of \emph{geometric quadrilaterals (quads)}, each one encoded as an array of quadruples of integer indices, pointing to the linear array of grid vertices associated to the standard image block $B$.  In other words, \emph{all quads} of \textbf{each job} are now given in the \textbf{same} \emph{local coordinates}.  Besides to put each partial surface $\mathcal{S}(B) = (\texttt{V}_B, \texttt{FV}_\sigma)$ in the global coordinate system of the image, the present stage must eliminate the redundant boundary features possibly generated at the edges of the partial surface $\mathcal{S}(B)$ within each block $B$ such that $B \cap \mathcal{I} \not= \emptyset$:
\begin{enumerate}

\item translate each array $\mathtt{FV}_\sigma$, of type \texttt{Lar.Cells}, by summing to each vertex index the linearized offset of the Cartesian coordinates $(n,m,p)(B)$ of the $B$'s \emph{reference vertex}, i.e.~the one with (all) lowest \emph{Cartesian coordinates} within the $B$ block.

@@

\item 
a smart strategy of removal of such artifacts may be used, which does not require any sorting nor searching on the assembled array of quads. It will consist in arranging each block with all three dimensions decreased-increased by one, so that each 2-adjacent pair of blocks will be covering each other for a full side extent of blocks of depth one. The details of this \emph{artifact filtering} will be elucidated in Section~\ref{sec:covering}.

@@

\subsubsection{Performance analysis}\label{sec:analysis}

>>>>

\section{Julia implementation}\label{sec:implementation}

@@

\subsection{Parallel workflow}\label{sec:implementation}

@@

\item depending on the above, best decision for $\emph{size}$ of 3D image blocks $\mathcal{B}$ (bricks); defaults to: $\emph{size}=64$; hence the number of bricks will be $n=\ceil{\ell_1/size}\times\ceil{\ell_2/size}\times\ceil{\ell_3/size}$. 
Hence default value: $n=256$, for images $\mathcal{I} = 512\times 512\times 256$;

\item computation of a Julia's sparse boundary matrix $[\partial_B]$, returning a value of type \texttt{SparseMatrixCSC\{Int8\}\{Int64\}}, where \texttt{Int8} and \texttt{Int64} are the types for values and indices, respectively, stored by Compressed Sparse Column (CSC) format; the average $[\partial_B]$ (for $\emph{size}$ = 64) is about 45 MB;

\item creation of either a local or distributed channel to implement a producer/consumer model of parallel/distributed computation, depending on available resources;

\item distribution of matrix $[\partial_B]$, of default size $45$ MB, to all available nodes/cores (workers) using the macro \texttt{@eveywhere}.

@@

The memory occupancy of the sparse matrix is computed by considering $8+1$ bytes for non-zero element (which are exactly 6 per row), plus 8 bytes per each index of column start.  
With $size=64$, the number of non-zeros within the sparse matrix $[\partial_{64^3}]$ is $\mathtt{nnz} = 4 792 266$, for a storage size of $9\times \mathtt{nnz}+8\times 262144 \simeq 45$ MB.

@@

\paragraph{3-Chain encoding} 
The interesting part of the \emph{Image} $\mathcal{I}$ is called \emph{Segment} $\mathcal{S}$. The goal of the whole \emph{workflow} is to extract a \emph{boundary model} of $\mathcal{S}$ from $\mathcal{I}$. The portion of $\mathcal{S}$ inside $\mathcal{B}$, will be denoted as $\mathcal{S}(B)$.
Each block $\mathcal{B}$ of the 3D image must by converted into the \emph{coordinate representation} of a vector $\nu\in C_3$ in the linear space  of 3-chains.

In coordinates local to $\mathcal{B}$, once fixed an ordering from Cartesian to linear coords, this vector is represented by a \emph{binary array} of length $size^3$. With $size=64$, we have $64^3=262144$,  with a non-zero value (i.e.~$1$) for each foreground voxel in $\mathcal{S}(B)$. Therefore, the coded segment portion $\mathcal{S}(B)$, results with a space occupancy of about $262$ KB if encoded as a full array (i.e.~including the zero values). Whether encoded as a sparse vector, its space occupancy will correspondingly decrease.

@@

\item conversion from linear indices to Cartesian indices in coordinates local to the $\mathcal{B}$ block, using the appropriate library functions;

@@

\paragraph{Assembling and artifact filtering} 
The results of the previous stages can be described as a \emph{collection of sets} of \emph{geometric quadrilaterals (quads)}, each one encoded as an array of quadruples of integer indices, pointing to the linear array of grid vertices associated to the image block $\mathcal{B}$.  In other words, \emph{all quads} of \textbf{each job} are now given in the \textbf{same} \emph{local coordinates}.  Besides to put each partial surface $\mathcal{S}(B) = (\texttt{V}_B, \texttt{FV}_\sigma)$ in the global coordinate system of the image, the present stage must eliminate the redundant boundary features possibly generated at the edges of the partial surface $\mathcal{S}(B)$ within each block $\mathcal{B}$ such that $B \cap \mathcal{I} \not= \emptyset$:
\begin{enumerate}

\item translate each array $\mathtt{FV}_\sigma$, of type \texttt{Lar.Cells}, by summing to each vertex index the linearized offset of the Cartesian coordinates $(n,m,p)(B)$ of the $\mathcal{B}$'s \emph{reference vertex}, i.e.~the one with (all) lowest \emph{Cartesian coordinates} within the $\mathcal{B}$ block.

@@

\item 
a smart strategy of removal of such artifacts may be used, which does not require any sorting nor searching on the assembled array of quads. It will consist in arranging each block with all three dimensions decreased-increased by one, so that each 2-adjacent pair of blocks will be covering each other for a full side extent of blocks of depth one. The details of this \emph{artifact filtering} are elucidated in Section~\ref{sec:covering}.

@@

\subsection{Performance analysis}\label{sec:analysis}

### 2019-06-30 src/main.tex
tables added
my last update

%SetFonts
\usepackage[T1]{fontenc}
\usepackage{lmodern}

@@

\title{Fast algebraic filtering of surfaces from 3D medical images with Julia}
\author{Miroslav Ji\v{r}\'ik and Alberto Paoluzzi}
\date{}							% Activate to display a given date or no date

@@

\begin{abstract}
In this paper we introduce a novel algebraic \textsc{lar-surf} filter, well founded on algebraic topology methods, to extract and smooth the boundary surface of any subset of voxels arising from the segmentation of a 3D medical image. The input is defined as a \emph{chain}, i.e.~as a vector from a linear space of 3-chains, represented in coordinates as a sparse Boolean vector. The output is produced as the result of the mapping via the linear boundary operator $\partial_3:C_3 \to C_2$ between linear spaces of 3- and 2-chains.  In particular, when the input set of voxels is either not (4-)connected, or contains one or more empty regions inside, \textsc{lar-surf} generates a non connected set of closed surfaces, i.e.~a set of 2-cycles---using the language of algebraic topology. The only data structures used by this approach are sparse arrays with one or two indices, i.e.~sparse vectors and matrices. This work is based on LAR (Linear Algebraic Representation) methods, and is implemented in Julia language, natively supporting parallel computing on hybrid architectures.
\end{abstract}

@@

\begin{description}
\item[$\mathcal{I}$]  three-dimensional medical image
\\[-8mm]
\item[$\ell_1, \ell_2, \ell_3$]  dimensions of image
\\[-8mm]
\item[$\mathcal{S}$]  segment: subset of voxels
\\[-8mm]
\item[$\mathcal{B}$] 3D image block 
\\[-8mm]
\item[$size$] lateral dimension of cubic block $\mathcal{B}$ 
\\[-8mm]
\item[$n$]	number of blocks $\mathcal{B}$ (jobs) in $\mathcal{I}$
\\[-8mm]
\item[[$\partial_\mathcal{B}$] ]  boundary matrix for block $\mathcal{B}$
\\[-8mm]
\item[$C_p$] linear (vector) space of $p$-chains
\\[-8mm]
\item[$\nu\in C_p$] $p$-chain
\\[-8mm]
\item[[$\nu$]] coordinate representation (binary vector) of  $\nu$
\\[-8mm]
\item[$\E^3$] Euclidean 3-space
\end{description}

\subsection{Definitions}
\begin{description}
\item[CSC]  Compressed Sparse Column format for sparse matrices
\\[-8mm]
\item[Global coordinates]  Integer linear coordinates of $\mathcal{I}$
\\[-8mm]
\item[Local coordinates] Integer linear coordinates of $\mathcal{B}$
\\[-8mm]
\item[Cartesian coordinates] Integer triples $(i,j,k)$ one-to-one with voxels 
\\[-8mm]
\item[Voxels] individual elements in 3D image (3-cells)
\\[-8mm]
\item[$p$-chain] formal linear combination of $p$-cells with coefficients in $\{0,1\}$
\\[-8mm]
\item[Coord. repr.]  binary vector (for $p$-chains) or binary matrix (for chain operators)
\\[-8mm]
\item[Quad]	geometric quadrilateral; convex polygon with four vertices 
\\[-8mm]
\item[Foreground voxel] individual element of a segment $\mathcal{S}$
\\[-8mm]
\item[Segment]	subset of voxels resulting from image segmentation
\end{description}

>>>>

%SetFonts
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

@@

\title{Algebraic filtering of surfaces from medical images with Julia}
\author{Miroslav Jirik and Alberto Paoluzzi}
\date{}							% Activate to display a given date or no date

@@

\begin{abstract}
In this paper we introduce a novel algebraic \textsc{lar-surf} filter, well founded on algebraic topology methods, to extract and smooth the boundary surface of any subset of voxels arising from  segmentation of a 3D medical image. The input is defined as a \emph{chain}, i.e.~as a vector from a linear space of 3-chains, represented in coordinates as a sparse binary vector. The output is produced by a linear mapping between spaces of 3- and 2-chains through the boundary operator $\partial_3:C_3 \to C_2$.  In particular, when the input set of voxels is either not (4-)connected, or contains one or more empty regions inside, \textsc{lar-surf} generates a non connected set of closed surfaces, i.e.~a set of 2-cycles---using the language of algebraic topology. The only data structures used by this approach are sparse arrays with one or two indices, i.e.~sparse vectors and matrices. This work is based on LAR (Linear Algebraic Representation) methods, and is implemented in Julia language, natively supporting parallel computing on hybrid architectures.
\end{abstract}

@@

\begin{description}
\item[$\mathcal{I}$]  three-dimensional medical image
\\[-7mm]
\item[$\ell_1, \ell_2, \ell_3$]  dimensions of image
\\[-7mm]
\item[$\mathcal{S}$]  segment: a subset of voxels from image segmentation
\\[-7mm]
\item[$\mathcal{B}$] 3D image block 
\\[-7mm]
\item[$size$] lateral dimension of cubic block $\mathcal{B}$ 
\\[-7mm]
\item[$n$]	number of blocks $\mathcal{B}$ (jobs) in $\mathcal{I}$
\\[-7mm]
\item[[$\partial_\mathcal{B}$] ]  boundary matrix for block $\mathcal{B}$
\\[-7mm]
\item[$C_p$] linear (vector) space of $p$-chains
\\[-7mm]
\item[$\nu\in C_p$] $p$-chain
\\[-7mm]
\item[[$\nu$]] coordinate representation (binary vector) of  $\nu$
\\[-7mm]
\item[$\E^3$] Euclidean 3-space
\end{description}

\subsection{Definitions}
\begin{description}

\item[Boundary model] Closed manifold surface of the boundary of a solid model
\\[-7mm]
\item[CSC]  Compressed Sparse Column format for sparse matrices
\\[-7mm]
\item[Global coordinates]  Integer linear coordinates of $\mathcal{I}$
\\[-7mm]
\item[Local coordinates] Integer linear coordinates of $\mathcal{B}$
\\[-7mm]
\item[Cartesian coordinates] Integer triples $(i,j,k)$ one-to-one with voxels 
\\[-7mm]
\item[Voxels] Individual elements in 3D image (3-cells)
\\[-7mm]
\item[$p$-chain] Formal linear combination of $p$-cells with coefficients in $\{0,1\}$
\\[-7mm]
\item[Coord. repr.]  Binary vector (for $p$-chains) or binary matrix (for chain operators)
\\[-7mm]
\item[Quad]	Geometric quadrilateral; convex polygon with four vertices 
\\[-7mm]
\item[Foreground voxel] Individual element of a segment $\mathcal{S}$
\\[-7mm]
\item[Segment]	Subset of voxels resulting from image segmentation
\end{description}

% Bibliography
\bibliographystyle{acm}
%\bibliography{TOMS2017.bib,proposal.bib,cadj2017.bib,bib.bib,schemes.bib}
\bibliography{TSAS-2019.bib}
